### 2016-09-16 GPCE2016/sections/Introduction.tex
spellcheck

Regardeless of the OO model, inheritance is a key mechanism is OO 
languages. Inheritance provides a modularization mechanism, which 
is used to reuse implementations from inherited classes/objects. 
Unfortunatelly, as widely acknowledged in the literature~\cite{scharli03traits,Sak89dis,bracha90mixin,malayeri2009cz}
multiple inheritance, especially when combined with state, has several
tricky issues, including several variants of the famous 
\emph{diamond problem}~\cite{bracha90mixin,Sak89dis}. Many of the problems related to 
inheritance arise from the direct use of \emph{fields} to model state. 
Inheriting two fields with the same name raises 
the question of whether the two fields should be kept, or only one 
field should exist. Initialization 
of the fields is also problematic, since initialization code may be inherited from 
multiple parents. Finally, an additional problem with mutable fields is that their 
type cannot be type-refined in extensions, which can cause modularity problems
~\cite{wadler98expression,eptrivially}.

>>>>

Regardless of the OO model, inheritance is a key mechanism is OO 
languages. Inheritance provides a modularization mechanism, which 
is used to reuse implementations from inherited classes/objects. 
Unfortunately, as widely acknowledged in the literature~\cite{scharli03traits,Sak89dis,bracha90mixin,malayeri2009cz}
multiple inheritance, especially when combined with state, has several
tricky issues, including several variants of the famous 
\emph{diamond problem}~\cite{bracha90mixin,Sak89dis}. Many of the problems related to 
inheritance arise from the direct use of \emph{fields} to model state. 
Inheriting two fields with the same name raises 
the question of whether the two fields should be kept, or only one 
field should exist. Initialization 
of the fields is also problematic, since initialization code may be inherited from 
multiple parents. Finally, an additional problem with mutable fields is that their 
type cannot be type-refined in extensions, which can cause modularity problems
~\cite{wadler98expression,eptrivially}.

### 2016-09-16 GPCE2016/sections/RelatedWork.tex
consistency.
spellcheck

\para{Traits vs Object Interfaces.}
We consider object interfaces an alternative to traits or mixins. In
trait model two concepts (traits and classes) coexist and cooperate. Some
authors~\cite{BettiniDSS13} see this as good language design fostering good
software development by helping programmers think about the program structures.
However, others see the need of two concepts and the
absence of state as drawbacks of this model~\cite{malayeri2009cz}.  Object
interfaces are units of reuse, and meanwhile provide factory methods
for instantiation and support state. Our approach promotes the use of
interfaces in order to exploit the modular composition
offered by interfaces. Since Java was designed for classes, a direct classless
programming style is verbose and unnatural. However, annotation-driven code
generation is enough to overcome this difficulty and the resulting programming
style encourages modularity, composability and reusability. %, by keeping a strong OO feel.
In that sense, we promote object interfaces as being both units of
reuse and object factories. Our practical experience shows that
separating the two notions leads to lots of boilerplate code, and is quite
limiting when multiple inheritance with state is required.  Abstract state
operations avoid the key difficulties associated with multiple inheritance and
state, while still being quite expressive. Moreover the ability to support
constructors adds expressivity, which is not available in approaches
such as Scala's traits/mixins.
%\footnote{Scala traits have
%  no constructor, but \Q@var@s and (\Q@lazy@) \Q@val@s can be
%  ``pre-initialized''.}

@@

\para{Automatic generation of getters and setters}
% \bruno{Need to discuss automatic generation of setter and
%   getters, and then discuss this text about injections. See reviews
%   and reply.}
is an old idea used in languages such as
Self~\cite{Ungar1987self}, Dart~\cite{Dart} and Newspeak~\cite{Bracha_thenewspeak}.
% By declaring fields in Newspeak and other languages,
The programmers specifies field signatures and (critically) the intention of storing such information,
then the language generates getters and setters.
Once state is abstracted away, it is well known that state
access can be replaced with computation, but the type of the field stays the same.
We do the opposite, the idea of the field is generated starting from signatures of
getters and setters. % Our fluid approach to state is specular:
In our approach, the intention of storing the information is not expressed by the programmer and set in stone
but can vary by inheritance.
In this case, the underling type of the field can be changed by our fluid state,
 and \Q@with@ methods provide the right injection from the old type to the new.

>>>>

\para{Traits vs object interfaces.}
We consider object interfaces an alternative to traits or mixins. In
trait model two concepts (traits and classes) coexist and cooperate. Some
authors~\cite{BettiniDSS13} see this as good language design fostering good
software development by helping programmers think about the program structures.
However, others see the need of two concepts and the
absence of state as drawbacks of this model~\cite{malayeri2009cz}.  Object
interfaces are units of reuse, and meanwhile provide factory methods
for instantiation and support state. Our approach promotes the use of
interfaces in order to exploit the modular composition
offered by interfaces. Since Java was designed for classes, a direct classless
programming style is verbose and unnatural. However, annotation-driven code
generation is enough to overcome this difficulty and the resulting programming
style encourages modularity, composability and reusability. %, by keeping a strong OO feel.
In that sense, we promote object interfaces as being both units of
reuse and object factories. Our practical experience shows that
separating the two notions leads to lots of boilerplate code, and is quite
limiting when multiple inheritance with state is required.  Abstract state
operations avoid the key difficulties associated with multiple inheritance and
state, while still being quite expressive. Moreover the ability to support
constructors adds expressivity, which is not available in approaches
such as Scala's traits/mixins.
%\footnote{Scala traits have
%  no constructor, but \Q@var@s and (\Q@lazy@) \Q@val@s can be
%  ``pre-initialized''.}

@@

\para{Automatic generation of getters and setters}
% \bruno{Need to discuss automatic generation of setter and
%   getters, and then discuss this text about injections. See reviews
%   and reply.}
is an old idea used in languages such as
Self~\cite{Ungar1987self}, Dart~\cite{Dart} and Newspeak~\cite{Bracha_thenewspeak}.
% By declaring fields in Newspeak and other languages,
The programmers specifies field signatures and (critically) the intention of storing such information,
then the language generates getters and setters.
Once state is abstracted away, it is well known that state
access can be replaced with computation, but the type of the field stays the same.
We do the opposite, the idea of the field is generated starting from signatures of
getters and setters. % Our fluid approach to state is specular:
In our approach, the intention of storing the information is not expressed by the programmer and set in stone
but can vary by inheritance.
In this case, the underlying type of the field can be changed by our fluid state,
 and \Q@with@ methods provide the right injection from the old type to the new.
